---
title: "Model comparison report"
author: "PTH"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

Report generated on `r Sys.time()`

```{r, message=FALSE, warning=FALSE, include=FALSE}
suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
suppressWarnings(suppressPackageStartupMessages(library(reshape2)))
suppressWarnings(suppressPackageStartupMessages(library(broom)))
suppressWarnings(suppressPackageStartupMessages(library(ggpubr)))
suppressWarnings(suppressPackageStartupMessages(library(xtable)))
options(xtable.comment = FALSE)
suppressWarnings(suppressPackageStartupMessages(library(brms)))
suppressWarnings(suppressPackageStartupMessages(library(broom)))
suppressWarnings(suppressPackageStartupMessages(library(bayesplot)))

library(knitr)
library(markdown)
library(rmarkdown)
```

# Dataset: `r paste0(SET)`

## Model comparisons with LOOIC

```{r results='asis', message=FALSE, warning=FALSE}
print(xtable(loo1))
```

Comparisons among top 5 models via LOO-IC:

```{r results='asis', message=FALSE, warning=FALSE}
print(xtable(tidy(loo_comp1[[6]])))
```

Best model is `r paste0(loo1$model_name[1])`. Examining this model's population-level coefficient estimates:

```{r results='asis', message=FALSE, warning=FALSE}
print(xtable(tidy(best_mod, par_type = "non-varying", prob = 0.95)))
```

Let's look at the coefficients for all the models:
```{r results='asis', message=FALSE, warning=FALSE}
print(xtable(tidy(h1a, par_type = "non-varying", prob = 0.95)))
```

Plotting posterior distributions of parameter estimates:

```{r fig.align='center', fig.height=3, fig.width=6.5, message=FALSE, warning=FALSE}
post_dists <- as.matrix(best_mod)
plot_title <- ggtitle("Posterior distributions",
                      "with medians and 95% intervals")

# grab relevant columns of post_dists
post_dists2 <- post_dists[,!(colnames(post_dists) %in% c(grep('r_',colnames(post_dists),value = T),grep('lp_',colnames(post_dists),value = T)))]
mcmc_areas(post_dists2,
#           regex_pars = c('b_','sd_','alpha'),
           prob = 0.95) + plot_title
```

### Posterior predictive model checking

Check 1: posterior predictive intervals by `herb_dmg` category:

```{r fig.height=2, fig.width=2, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center'}
yrep = posterior_predict(best_mod)
# just do simple PPC plot using brms built-in functions, for now:
if(1 %in% grep("0", loo1$model_name)){
  ppc_plot1 <- ppc_dens_overlay(
  y = best_mod$data$log_ratio,
  yrep = yrep[1:100,],
  col = "gray80")
} else {
  ppc_plot1 <- ppc_violin_grouped(
  y = best_mod$data$log_ratio,
  yrep = yrep,
  #x = as.numeric(paste0(best_mod$data$herb_dmg)),
  group = best_mod$data$herb_dmg,
  #prob = 0.95,
  col = "gray80")
}
print(ppc_plot1)
```

## Predicting CFU counts from best log-ratio model `r paste0(loo1$model_name[1])`

First let's examine the CFU model used to generate the predictions, for book-keeping:

```{r results='asis'}
#print(xtable(tidy(cfu_mod, par_type = "non-varying")))
print(xtable(tidy(cfu_mod)))
```

Now let's output summaries of total log mean and median CFU per leaf class:

```{r results='asis'}
print(xtable(all_sum[,grep('mu',names(all_sum))]))
```

```{r results='asis'}
print(xtable(all_sum[,grep('med',names(all_sum))]))
```

Plots of distributions of predicted sample-level log-ratios and the resulting log predicted CFU per leaf class, Family-level, for rep `r paste0(the_rep)`:

```{r fig.height=2, fig.width=3, message=FALSE, warning=FALSE, paged.print=FALSE, align='center', fig.cap='Distributions of posterior predicted log-ratios (left) and log CFU (right) for a random replicate simulation.'}
# need to produce these plots:
ggarrange(plotlist = list(lr_raw, plot_raw1), ncol = 2)
```

Plots of distributions of medians and means of the log posterior predicted CFU per leaf class:

```{r fig.height=2, fig.width=6, message=FALSE, warning=FALSE, paged.print=FALSE, fig.align='center', fig.cap='Distributions of posterior predicted median log CFU (right) and mean log CFU (middle, right) across 1000 posterior simulation replicates.'}
ggarrange(plotlist = list(all_plot1_med, all_plot1, all_plot2, all_plot2_med), ncol = 4)
```

