---
title: "OTU model report"
author: "PTH"
date: "October 12, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
suppressWarnings(suppressPackageStartupMessages(library(dplyr)))
suppressWarnings(suppressPackageStartupMessages(library(reshape2)))
suppressWarnings(suppressPackageStartupMessages(library(broom)))
suppressWarnings(suppressPackageStartupMessages(library(gridExtra)))
suppressWarnings(suppressPackageStartupMessages(library(xtable)))
suppressWarnings(suppressPackageStartupMessages(library(brms)))
suppressWarnings(suppressPackageStartupMessages(library(bayesplot)))
library(knitr)
library(markdown)
library(rmarkdown)
```
## Basic statistics for OTU `r otu_name`
Point estimated based on the data for prevalence and abundance in both datasets is listed below:
```{r}
char_stop <- which(match(unlist(strsplit(otu_name,'')),'_', F)==1)-1
#this_otu_stat <- otu_stats[grep(substr(otu_name, start = 0, stop = char_stop), otu_stats[,'taxon_1perc']),]
# need to remove grep.. matches multiple OTUs!!
if (length(char_stop > 1)){
  this_otu_stat <- dplyr::filter(otu_stats, taxon_1perc == substr(otu_name, start = char_stop[1]+2, stop = char_stop[2]))
  this_otu_name <- paste(this_otu_stat[,'taxon_1perc'])  
} else {
  this_otu_stat <- dplyr::filter(otu_stats, taxon_1perc == substr(otu_name, start = 0, stop = char_stop))
  this_otu_name <- paste(this_otu_stat[,'taxon_1perc'])  
}

print(this_otu_stat)
```

Looking more closely at the dataset `r substr(otu_tab[1,'samples'], start = 1, stop = 4)`, we have empirical point estimates of prevalence and abundance broken down by herbivory as follows. 

```{r}
N <- length(otu_tab[,this_otu_name])
Ndmg <- sum(otu_tab[,'herb.dmg'])
Nun <- N - Ndmg
tots <- c(Nun,Ndmg,N)
prev_estimates <- data.frame(rbind('counts' = c(sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'herb.dmg']==0]),
                                                sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'herb.dmg']==1]),
                                                sum(as.numeric(otu_tab[,this_otu_name]>0))),
                                  'totals' = round(tots),
                                  'freqs' = round(c(sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'herb.dmg']==0]),
                                                    sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'herb.dmg']==1]),
                                                    sum(as.numeric(otu_tab[,this_otu_name]>0))) / tots, 3)))
names(prev_estimates) <- c('undamaged','damaged','total')
```

```{r results='asis', message=FALSE, warning=FALSE}
print(xtable(prev_estimates))
```

The empirical relative abundance distribution of counts is displayed below (includes zeros) for both herbivory classes:

```{r, fig.align='center', fig.height=2.5, fig.width=4.5, message=FALSE, warning=FALSE}
relabund <- data.frame(samples = otu_tab[,'samples'],
                       herb.dmg = otu_tab[,'herb.dmg'],
                       stem.tx = otu_tab[,'stem.tx'],
                       relabund = otu_tab[,this_otu_name]/otu_tab[,'total'])
ggplot(relabund, aes(x = relabund)) + geom_histogram() + facet_wrap(~ herb.dmg) #+ scale_x_log10()
```

### Hormone treatment prevalences
The following table displays the prevalence of this OTU across each of the hormone treatment categories.

```{r}
tots2 <- c(table(otu_tab[,'stem.tx']),'total' = N)
the_counts <- c(sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'stem.tx']=='MC']),
                                     sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'stem.tx']=='JATX']),
                                     sum(as.numeric(otu_tab[,this_otu_name]>0)[otu_tab[,'stem.tx']=='SATX']),
                                     sum(as.numeric(otu_tab[,this_otu_name]>0)))
prev_estimates2 <- data.frame(rbind('counts' = the_counts,
      'totals' = tots2,
      'freqs' = round(the_counts/tots2, 3)))
names(prev_estimates2) <- c('MC','JATX','SATX','total')
```


```{r results='asis', message=FALSE, warning=FALSE}
print(xtable(prev_estimates2))
```


Now a plot of relative abundances in samples across hormone treatment types:

```{r, fig.height=2.5, fig.width=6, fig.align='center', message=FALSE, warning=FALSE}
ggplot(relabund, aes(x = relabund)) + geom_histogram() + facet_wrap(~ stem.tx)
```

## Regression model report for `r otu_name`
### Summary of model output
```{r}
summary(OTU, waic=T)
```

The LOOIC for this model is below:
```{r, echo=FALSE, message=FALSE, warning=FALSE}
LOO(OTU)
```

Coefficient estimates (95% posterior intervals) are depicted below:

```{r, echo=FALSE, fig.height=2.5, fig.width=5.5, message=FALSE, warning=FALSE}
stanplot(OTU)
```

Based on these model results, the 2D reaction norm plot is the following:
```{r, message=FALSE, warning=FALSE, include=FALSE}
# calculate prevalences in both damage classes
coefs <- tidy(OTU)[1:8,]

prevs <- 1 - exp(coefs[c(5,6),'estimate'])/(1+exp(coefs[c(5,6),'estimate']))
prevs_H <- 1 - exp(coefs[c(5,6),'lower'])/(1+exp(coefs[c(5,6),'lower']))
prevs_L <- 1 - exp(coefs[c(5,6),'upper'])/(1+exp(coefs[c(5,6),'upper']))

# Rates
# abunds <- exp(coefs[c(1,2),'estimate'])
# abunds_L <- exp(coefs[c(1,2),'lower'])
# abunds_H <- exp(coefs[c(1,2),'upper'])

# log rates
abunds <- coefs[c(1,2),'estimate']
abunds_L <- coefs[c(1,2),'lower']
abunds_H <- coefs[c(1,2),'upper']

the_res <- data.frame(prevs, prevs_L, prevs_H, abunds, abunds_L, abunds_H, herb.dmg = c(0,1))
write.table(the_res, file = paste0(OUT_DIR,this_otu_name,"_model_coefs_calculated.txt"), row.names = FALSE, quote = F, sep = '\t')
#print(round(the_res,3))
```

Below is a two dimensional reaction norm plot based on 95% credible intervals of posterior estimates for prevalence and a measure of abundance (conditional on presence).

```{r, echo=FALSE, fig.align='center', fig.height=4, fig.width=4, message=FALSE, warning=FALSE}
# make plot of 2D reaction norm with and without herbivory
ggplot(the_res,aes(x = prevs, y = abunds, xmin = prevs_L, xmax = prevs_H, ymin = abunds_L, ymax = abunds_H)) + geom_path(aes(x = prevs, y = abunds)) +
  geom_errorbar(width = 0, col = "gray60") + geom_errorbarh(height = 0, col = "gray60") +
  geom_point(aes(col = factor(herb.dmg)), size = 4) + 
  scale_color_manual("damage", values = c("black","darkorange2")) + scale_x_continuous(limits = c(0,1)) + #scale_y_continuous(limits = c(-10,0)) +
  theme(legend.position = "bottom") + xlab("prevalence") + ylab("abundance (nb rate)") 
  #theme_bw()
```

## Posterior predictive checks of the model.

```{r, echo=FALSE, fig.align='center', fig.height=5, fig.width=5, message=FALSE, warning=FALSE}
y <- otu_tab[,this_otu_name]
yrep <- posterior_predict(OTU, nsamples = 500)

if (sum(is.na(yrep)) == 0){
  
# break y and yrep down by herb.dmg:
y1 <- y[otu_tab[,'herb.dmg']==0]
yrep1 <- yrep[,otu_tab[,'herb.dmg']==0]

y2 <- y[otu_tab[,'herb.dmg']==1]
yrep2 <- yrep[,otu_tab[,'herb.dmg']==1]

ppc1 <- ppc_dens_overlay(log(1+y1), log(1+yrep1)) + theme(legend.position = "top") + ggtitle("(a) undamaged")
ppc2 <- ppc_dens_overlay(log(1+y2), log(1+yrep2)) + theme(legend.position = "top") + ggtitle("(b) damaged")
grid.arrange(ppc1,ppc2)
} else{
  print("Model did not converge; check Rhat values. No posterior predictive check was run!")
}
```


